Unsupervised Acquisition of Comprehensive Multiword Lexicons \\ using Competition in an $n$-gram Lattice

### 2017-01-19 tacl.tex
Switched from cost to expainedness

Typically multiword lexicons are created by ranking $n$-grams according to an association measure and applying a threshold. The algorithm of Silva and Lopes \shortcite{Silva99} is somewhat more sophisticated, in that it identifies the local maxima of association measures across subsuming $n$-grams within a sentence to identify MWEs of unrestricted length; its effectiveness beyond noun phrases, however, appears limited \cite{Ramisch12}. Brooke et \al \shortcite{Brooke14a,Brooke15b} developed a heuristic method intended for general FS extraction in larger corpora, first using conditional probably statistics to do an initial (single pass) coarse-grained segmentation of the corpus, followed by a pass through the resulting vocabulary, breaking larger units into smaller ones based on tradeoff between marginal and conditional statistics (\ie LPR). Beyond association measures, other general unsupervised approaches to the multiword unit identification include that of Newman et \al \shortcite{Newman12}, who used a generative Dirichlet Process model which jointly creates a linear segmentation of the corpus and a multiword vocabulary of keyphrases. Gimpel and Smith \shortcite{Gimpel11} focus specifically on deriving word sequences with gaps using a generative model, with the intent of improving machine translation. A general drawback to these generative methods, relative to approaching involving association metrics, is scalability and lack of determinacy.

@@

Our approach to FS identification involves optimization of the total cost of a lattice where each node corresponds to an $n$-gram type. The cost of the whole lattice is defined simply as a product of the costs of the individual nodes. Each node can be considered either ``on'' (is an FS) or ``off'' (is not an FS), and incur different costs depending on their status. Most fundamentally, $n$-gram nodes in the lattice are directionally connected to nodes consisting of $n+1$-grams which subsume them and $n-1$-grams which they subsume: that is, the (gapped) $n$-gram \ex{keep \gap under wraps} would, for instance be connected ``upwards'' to the node \ex{keep everything under wraps} and connected ``downwards'' to \ex{under wraps}. These directional relationships allow for two basic interactions between nodes in the lattice when a node is turned on: \termdef{covering}, which inhibits nodes below (subsumed by) a turned-on node (\eg, if \ex{keep \gap under wraps} is on, the model will tend not to choose \ex{under wraps} as an FS); and \termdef{clearing}, which inhibits nodes above a turned-on node (\eg, if \ex{keep \gap under wraps} is on, the model would avoid selecting \ex{keep everything under wraps} as an FS). A third, undirected mechanism is \termdef{overlapping}, where nodes inhibit each other due to overlaps in the corpus (\eg having both \ex{keep \gap under wraps} and \ex{be keep \gap under} as FS will be avoided). Given these relationships, we minimize the cost of the entire lattice by iterating over the nodes and greedily optimizing the on/off choice relative to the local neighborhood of each node, until convergence.

@@

In the segmentation approach of Brooke et \al \shortcite{Brooke15b}, LPR for an entire span is calculated as a product of the individual LPRs, but here the base ``off'' cost for a node in our lattice is the minimum LPR across the words in the sequence (minLPR):

@@

\subsection{Node cost functions}

The central process of the model is to decide which nodes in an $n$-gram lattice correspond to an FS, based on the objective of minimizing the overall cost of the lattice. It is useful to define the cost of a node in terms of two cost functions. The rationale is that turning on a node completely changes the influence that surrounding nodes will have: when a node is off, it is generally desirable (or at least not undesirable) that connected nodes in the lattice are on, since they may serve to explain its existence. When a node is on, however, the presence of other turned-on nodes nearby suggests redundancy, and is not to be encouraged. Let $C_{1}$ be the cost of a node when it is on, corresponding to an $n$-gram which has been identified as an FS, and $C_{0}$ be the cost of a node when it off. The base value of $C_{0}$ is the minLPR metric discussed above. The base value of $C_{1}$, which we will call just $C$, is a parameter to the model which, for nodes not influenced by one of the other factors discussed below, is equivalent to an initial threshold for building a vocabulary using the minLPR association measure: all else being equal, when $minLPR(t) > C$, a node will be selected as an FS. There is no necessary upper bound on $C$, but we note here that there is a fairly small range of good choices for $C$; we have found that setting it too high ($>6$, corresponding to minimum of six times more likely than predicted by POS context alone) will restrict the model primarily to sequences with only open-class words, while setting it too low ($<3$) allows in too much random noise.

@@

We will influence these cost functions by adding exponential terms, reflecting the fact we do not want to override the association measure, but rather modulate it. Let's define $C_0$, then, as the minLPR based taken to the power of some as yet undefined function $d_0$:

\begin{displaymath}
C_0(t) = minLPR(t)^{d_0(t)}
\end{displaymath}
\noident

And $C_1$ is the cost parameter $C$ taken to the power of an analogous function $d_1$:

\begin{displaymath}
C_1(t) = C^{d_1(t)}
\end{displaymath}
\noident

The values of $d_0$ and $d_1$ are dependent on the context of the node in the lattice, but are both 1 when a node is not being influenced by other nodes. They are also restricted to always being positive: as such, the lower bound on both costs is 1.

@@

\termdef{Soft covering} is used in cases when a single $n$-gram does not entirely account for another, but a turned-on $n$-gram to some extent may explain some of the statistical irregularity of one lower in the lattice. For instance, in \figref{fig:example} \ex{keep \gap under} is not hard-covered by \ex{keep \gap under wraps} (since there are FS such as \ex{keep \gap under surveillance}, \ex{keep it under your hat}, \etc), but if \ex{keep \gap under wraps} is tagged as an FS, we nevertheless want to discount the portion of the \ex{keep \gap under} counts that corresponds to occurrences of \ex{keep \gap under wraps}. This is accomplished by lowering the turned-off cost function of \ex{keep \gap under} (and thus making turning on less desirable) in the following manner: let $c(\dot)$ be the count function,and $s_1\dotts s_m$ be any turned-on nodes which are above $t$ in the lattice (covering nodes). Then, then the $d_0$ for a covered node $t$ is:

\begin{displaymath}
d_{0}(t) = \frac{c(t) - \sum_{i=1}^{m}{c(s_i)}}{c(t)}
\end{displaymath}

@@

In general, covering prefers turning on longer, covering $n$-grams since doing so lowers the turned-off cost function of nodes lower in the lattice. Not surprisingly, it is generally desirable to have a mechanism working in opposition, \ie one which generally prefers shorter $n$-grams. \termdef{Clearing} does this by changing the $C_0$ of nodes higher in the lattice when a lower node is turned-on. The basic mechanism is similar to covering, except that counts cannot be made use of in the same way; whereas it makes sense to discount the turned-off cost of covered nodes in proportion to the counts of their covering nodes (since the counts of the covered $n$-grams are due to the covering $n$-gram), in the reverse direction this logic fails. A simple but effective solution which avoids introducing parameters makes use of the minLPR values of the relevant nodes. In the most common (two node) situation, we lower the  $C_{0}(t)$ of the higher node according to the ratio between the minLPR of two nodes, though only if the minLPR of the lower node is higher. Generalized to consider the (rare) case of multiple clearing nodes, this is:

@@

\begin{displaymath}
C_{0}'(t) = max(1, C_{0}(t) \frac{minLPR(t)}{\prod{minLPR(s_i)})}
\end{displaymath}
where $s_i$ is the $i$th clearing node. We refer to this mechanism as clearing because it tends to clear away a variety of trivial uses of common FS which many have higher LPR due to the lexical and syntactic specificity of the FS. To give an example of its effect, if the node \ex{keep \gap under wraps} in \figref{fig:example} is turned on and has a minLPR of 8, then, if the minLPR of a node such as \ex{keep \gap under wraps for} is 4, the $C_{0}$ will be lowered to 2 ($4/(8/4)$) by clearing. This dissuades (but does not prevent) the model from turning on these nodes. Note that the effect of both covering and clearing is to lower the cost function for covered/clears nodes, which encourages the model to turn an $n$-gram like \ex{keep \gap under wraps} on, even supposing its minLPR were relatively low.

@@

The third mechanism of node interaction involves $n$-grams which overlap in the corpus. Though there are some exceptions, it is generally the case that independent formulaic sequences do not overlap. For example, given that \ex{be keep \gap under} and \ex{keep \gap under wraps} often appear together, overlapping on the tokens \ex{keep \gap under}, we do not want both being selected as an FS, even in the case where both had high minLPR. To address this problem, we use a mechanism somewhat different than above: rather than lowering the off-cost, we raise the on-cost. The idea behind lowering the cost when covering and clearing is to encourage the model to turn on nodes which effectively explain the presence of nodes above and/or below it in the lattice, but we do not want to encourage nodes that just happen to overlap with others; a penalty is a more appropriate mechanism. Let $oc(x,y)$ refer to number of times the $n$-grams corresponding to nodes $x$ and $y$ overlapped in the corpus, and $o_1\dotts o_m$ refer to \termemph{turned-on} nodes which overlap with our target node $t$. We define an $d_1$ exponential term for the $C_1$ our target node $t$ as:

@@

The effect of overlaps on the cost is hyperbolic: small amounts of overlapping has little effect, but with significant overlap the cost becomes prohibitive, effectively forcing one of nodes to turn off.

@@

The dependence of the cost functions of nodes on their neighbors as discussed in the preceding section effectively prohibit a global optimization of the lattice. Fortunately, though a majority of nodes in the lattice are connected to each other though some path, most of the effects of nodes on each other are relatively local, and effective local optimizations can be made tractable by applying some simple restrictions. The main optimization loop consists of iterations over the lattice until complete convergence (no changes in the final iteration). For each iteration over the main loop, each potentially active node is examined in order to evaluate whether its current status is optimal given the current state of the lattice. The order that we do this has an effect on the result: among the obvious options, good results are obtained when ordering nodes by frequency, which gives an implicit advantage to relatively common $n$-grams, since they will tend to be turned on first.

@@

Given the relationships between nodes that we have defined above, it is obviously not sufficient to consider switching only the present node; If, for instance, one or more of \ex{be keep \gap under wraps}, \ex{under wraps}, or \ex{be keep \gap under} has been turned on, the covering, blocking, or overlapping effects of these other nodes will likely prevent a competing node like \ex{keep \gap under wraps} from being correctly activated. Instead, the algorithm identifies a small set of nodes (relevant nodes) which are the most important to status of the node under consideration whose status should be considered simultaneously. Since turned-off nodes have no direct effect on each other, only turned-on nodes above, below, or overlapping with the current node need be considered.  Once the relevant nodes have been identified, all nodes (including turned-off nodes) whose cost function is affected by one or more of the relevant nodes are identified, and then a greedy search is carried out for the optimal configuration of the relevant nodes starting from an `all-on' state.

@@

\begin{itemize}
\item The complexity of the algorithm increases exponentially with the number of relevant nodes, so we limit the total number to 5. When there are more than 5 nodes turned on in the vicinity of the target node, relevant nodes are selected by ranking in terms of the influence of the nodes on each other as given by the relative change in their cost functions.
\item The full set of pairwise $n$-gram overlaps in a large corpus is too big too effectively store, and so we exclude overlaps with a count of less than 5 from our lattice.
\item Many nodes have a minLPR which is a little larger than 1. There is very little chance these nodes will be activated by the algorithm, and so after applying hard covering we do not consider activating nodes with a minLPR of less than 2.
\end{itemize}

@@

	Lattice \emph{-ovr} & 0.52& \bf{0.53} & 0.53 & & 0.33& \bf{0.60} & 0.43& & 0.35 & \bf{0.65} & 0.45 & & 0.69 & \bf{0.51} & \bf{0.59} \\

				Lattice & \bf{0.67} & 0.45 & \bf{0.54} & & \bf{0.41}& 0.56 & \bf{0.47} & &\bf{0.42} & 0.57 & \bf{0.48} & & \bf{0.86} & 0.39 & 0.54 \\ 
            \hline
           \hline

@@

       \hline
			& P & R & F\\
			 \hline
			LocalMaxs & \bf{0.58} & 0.14 & 0.23 \\
			minLPRrank & 0.34 & 0.45 & 0.39 \\
			LPR-seg & 0.42 & 0.45 & 0.43 \\
			DP-seg & 0.35 & \bf{0.71} & 0.47 \\
			Lattice & 0.46 & 0.61 & \bf{0.53} \\
       \hline
 \end{tabular}

@@

Our primary comparison is with the heuristic LPR model from Brooke et \al \shortcite{Brooke15b}. For the BNC, we also compare, separately, the DP-seg model from Newman et \al \shortcite{Newman12}, and the LocalMaxs algorithm of Silva and Lopes \shortcite{Silva99}; because these other approaches only generate sequential multiword units, we use only the sequential part of the BNC test set. All these approaches have been compared against a wide range of association measures;\footnote{One popular association measure which is theoretically extensible to $n$-grams of any length but which has not been addressed in the context of unrestricted MWE/FS identification is the (log-)likelihood ratio \cite{Dunning93}. Calculating a likelihood ratio in the general case involves considering all possible ways an $n$-gram could have been generated statistically from its component words: that is, whereas minLPR only considers the addition of one word to each $n-1$-gram, likelihood ratios must consider all combinations of all component $n-1,2,3\ldots$-grams, including, in our case, not only contiguous but also non-contiguous components. The corresponding exponential increase in complexity renders it and other such methods impractical for unrestricted FS identification.} we do not repeat all these comparisons here, but we do consider a lexicon built from ranking $n$-grams according to the measure used in our lattice (minLPR) as well as PMI. For each of these two association measures we build a lexicon equal to the size of the lexicon produced by our model. All comparisons here are subject to the same $n$-gram frequency threshold as the main model, since the content of our test sets are also linked to this threshold.

@@

The main results for FS acquisition across all four corpora are shown in \tabref{tab:main}. As noted in previous work, simple statistical association measures like PMI do fairly poorly when faced with unrestricted $n$-grams of variable length: minLPR is clearly a much better statistic for this purpose. The LPRseg method of Brooke et \al \shortcite{Brooke15b} consistently outperforms simple ranking, and the lattice method proposed here does better still, with a margin that is fairly consistent across languages. Generally, using clearing and overlap node interactions offer an relatively large increase in precision at the cost of a smaller drop in recall, though the change is fairly symmetrical with in Croatian. The ICWSM corpus also has relatively high precision, whereas both the BNC and Croatian corpus have low precision and high recall.

In the continuous FS test set for the BNC (\tabref{tab:BNC}), we found that the LocalMaxs algorithm did fairly poorly, whereas the DP-seg method of Newman \cite{Newman12} was also able to beat our baseline methods, but with regards to F-score, it was inferior to our Lattice method. Some of that difference seems attributable to a severe precision/recall imbalance, though it was one we unable to improve by changing the parameters of the model from recommended settings.

@@

Though the results across the four corpora are reasonably similar with respect to overall f-score, there are a significant discrepancies. By using the UniDict morpheme representation as the starting unit for Japanese, the model is, in essence, doing an extra layer of FS identification, one which is provided by word boundaries in the other languages. The result is that there are many more FS for Japanese. Precision is very high, and recall is relatively low. Under these conditions, methods which tend to increase precision at the cost of recall will be less desirable with respect to f-score, which explains why the full model is not the best in that one case.  Importantly, the initial $n$-gram statistics derived from corpus reflect that Japanese is different: the number of $n$-gram types, over length 4, for instance, is almost twice the number in the ICWSM, or almost 14\% of all $n$-gram types. One idea for future work is to automatically adapt to the particular properties of the input language/corpus, though it might also be worth considering whether f-score is the most appropriate metric for evaluation, given that high precision seems particularly desirable in a lexical resource.

At the opposite end of the precision/recall curve, the low precision of the BNC is almost certainly at least somewhat a reflection of its relatively low size: whereas the $n$-gram threshold we have used here results in a minimum counts of in the order of about 100 for the other three corpora used here, the BNC statistics includes $n$-grams which appear less than 10 times. This might be resolved, of course, by increasing the $n$-gram threshold, or simply avoiding small corpora altogether, but it may be useful to focus effort on how to building comprehensive FS lexicons even in relatively low resource situations. One idea we are actively pursing is modifying the calculation of the LPR metric to integrating uncertainty due to low counts.

There is one another more general explanation for the precision/recall imbalances we see across the four corpora: Both the BNC and the Croatian corpus are composed primarily (though not exclusively) of published texts written by professional writers, in particularly newswire. The ICWSM and the Japanese corpus, on the other hand, are mostly blogs. Though all genres have FS, there is much more diversity in less controlled genres: many English gapped expressions, in particular, are used almost exclusively in relatively informal genres. The fact that last than 20\% of the FS derived from the BNC appear in the lexicon for the ICWSM is fairly telling of how important genre differences are to the range of FS that will be seen in a particular corpus.

>>>>

Typically multiword lexicons are created by ranking $n$-grams according to an association measure and applying a threshold. The algorithm of Silva and Lopes \shortcite{Silva99} is somewhat more sophisticated, in that it identifies the local maxima of association measures across subsuming $n$-grams within a sentence to identify MWEs of unrestricted length; its effectiveness beyond noun phrases, however, seems relatively limited \cite{Ramisch12}. Brooke et \al \shortcite{Brooke14a,Brooke15b} developed a heuristic method intended for general FS extraction in larger corpora, first using conditional probably statistics to do an initial (single pass) coarse-grained segmentation of the corpus, followed by a pass through the resulting vocabulary, breaking larger units into smaller ones based on tradeoff between marginal and conditional statistics (\ie LPR). Beyond association measures, other general unsupervised approaches to the multiword unit identification include that of Newman et \al \shortcite{Newman12}, who used a generative Dirichlet Process model which jointly creates a linear segmentation of the corpus and a multiword vocabulary of keyphrases. Gimpel and Smith \shortcite{Gimpel11} focus specifically on deriving word sequences with gaps using a generative model, with the intent of improving machine translation. A general drawback to these generative methods, relative to approaching involving association metrics, is scalability and lack of determinacy.

@@

Our approach to FS identification involves optimization of the total explanatory power of a lattice where each node corresponds to an $n$-gram type. The explanatory power of the whole lattice is defined simply as a product of the \termdef{explainedness} of the individual nodes. Each node can be considered either ``on'' (is an FS) or ``off'' (is not an FS), and explainedness will be calculated differently depending on their status. Most fundamentally, $n$-gram nodes in the lattice are directionally connected to nodes consisting of $n+1$-grams which subsume them and $n-1$-grams which they subsume: that is, the (gapped) $n$-gram \ex{keep \gap under wraps} would, for instance be connected ``upwards'' to the node \ex{keep everything under wraps} and connected ``downwards'' to \ex{under wraps}. These directional relationships allow for two basic interactions between nodes in the lattice when a node is turned on: \termdef{covering}, which inhibits nodes below (subsumed by) a turned-on node (\eg, if \ex{keep \gap under wraps} is on, the model will tend not to choose \ex{under wraps} as an FS); and \termdef{clearing}, which inhibits nodes above a turned-on node (\eg, if \ex{keep \gap under wraps} is on, the model would avoid selecting \ex{keep everything under wraps} as an FS). A third, undirected mechanism is \termdef{overlapping}, where nodes inhibit each other due to overlaps in the corpus (\eg having both \ex{keep \gap under wraps} and \ex{be keep \gap under} as FS will be avoided). Given these relationships, we maximize the explanatory power of the entire lattice by iterating over the nodes and greedily optimizing the on/off choice relative to the local neighborhood of each node, until convergence.

@@

In the segmentation approach of Brooke et \al \shortcite{Brooke15b}, LPR for an entire span is calculated as a product of the individual LPRs, but here we will use the minimum LPR across the words in the sequence (minLPR):

@@

\subsection{Node explainedness functions}

The central process of the model is to decide which nodes in an $n$-gram lattice correspond to an FS, based on the objective of minimizing the overall explanatory power of the lattice. Like probabilities, explainedness of individual nodes will vary between 0 and 1, with 1 corresponding to a node which is fully explained. It will be useful to define the explainedness of a node in terms of two functions: the rationale for having two functions is that the on/off status of a node completely changes the influence that surrounding nodes will have: when a node is off, it is generally desirable (or at least not undesirable) that connected nodes in the lattice are on, since they may serve to partially or entirely explain its existence. When a node is on, however, the presence of other turned-on nodes nearby suggests redundancy, and is not to be encouraged. Let $E_{1}$ be the explainedness of a node when it is on, corresponding to an $n$-gram which has been identified as an FS, and $E_{0}$ be the explainedness of a node when it off. The base value of $E_{0}$ is the inverse of the minLPR metric discussed above: when the minLPR metric is 1, corresponding to a case when the syntactic content predicts one of the words of the $n$-gram as well or better than the lexical context, $E_{0}$  will also be 1. The base value of $E_{1}$ is the inverse of a cost parameter $C$ and, for nodes not influenced by one of the other factors discussed below, is equivalent to the initial threshold for building a vocabulary using the minLPR association measure: all else being equal, when $minLPR(t) > C$, a node will be selected as an FS. There is no necessary upper bound on $C$, but we note here that there is a fairly small range of good choices for $C$; we have found that setting it too high ($>6$, corresponding to minimum of six times more likely than predicted by POS context alone) will restrict the model primarily to sequences with only open-class words, while setting it too low ($<3$) allows in too much random noise.

@@

We will influence these functions by adding exponential terms, reflecting the fact we do not want to override the association measure, but rather modulate it. Let's define $E_0$, then, as the minLPR based taken to the negative power of some as yet undefined function $d_0$:

\begin{displaymath}
E_0(t) = minLPR(t)^{-d_0(t)}
\end{displaymath}
\noident

And $E_1$ is the cost parameter $C$ taken to the negative power of an analogous function $d_1$:

\begin{displaymath}
E_1(t) = C^{-d_1(t)}
\end{displaymath}
\noident

The values of $d_0$ and $d_1$ are dependent on the context of the node in the lattice, but are both 1 when a node is not being influenced by other nodes. They are also restricted to always being positive.

@@

\termdef{Soft covering} is used in cases when a single $n$-gram does not entirely account for another, but a turned-on $n$-gram to some extent may explain some of the statistical irregularity of one lower in the lattice. For instance, in \figref{fig:example} \ex{keep \gap under} is not hard-covered by \ex{keep \gap under wraps} (since there are FS such as \ex{keep \gap under surveillance}, \ex{keep it under your hat}, \etc), but if \ex{keep \gap under wraps} is tagged as an FS, we nevertheless want to discount the portion of the \ex{keep \gap under} counts that corresponds to occurrences of \ex{keep \gap under wraps}. This is accomplished by increasing the turned-off explainedness of \ex{keep \gap under} (and thus making turning on less desirable) in the following manner: let $c(\dot)$ be the count function,and $s_1\dotts s_m$ be any turned-on nodes which are above $t$ in the lattice (covering nodes). Then, then the $d_0$ for a covered node $t$ is:

\begin{displaymath}
cover(t) = max(0,\frac{c(t) - \sum_{i=1}^{m}{c(s_i)}}{c(t)})
\end{displaymath}

@@

In general, covering prefers turning on longer, covering $n$-grams since doing so explains nodes lower in the lattice. Not surprisingly, it is generally desirable to have a mechanism working in opposition, \ie one which views shorter FS as helping to explain the presence of longer ones which contain them, beyond the FS-neutral syntactic explanation provided by minLPR. \termdef{Clearing} does this by changing the $E_0$ of nodes higher in the lattice when a lower node is turned-on. The basic mechanism is similar to covering, except that counts cannot be made use of in the same way; whereas it makes sense to explain covered nodes in proportion to the counts of their covering nodes (since the counts of the covered $n$-grams are from the covering $n$-gram), in the reverse direction this logic fails. A simple but effective solution which avoids introducing parameters makes use of the minLPR values of the relevant nodes. In the most common (two node) situation, we increase $E_{0}$ of the higher node based on the ratio of the minLPR of two nodes, though only if the minLPR of the lower node is higher, which greatly restricts its effect. Generalized to consider the (rare) case of multiple clearing nodes, we define clear(t) as:

@@

\begin{displaymath}
clear(t) = \prod_{i=1}^{m}{ min(1,\frac{minLPR(t)}{minLPR(u_i)})}
\end{displaymath}
where $u_i$ is the $i$th clearing node of $t$. We refer to this mechanism as clearing because it tends to clear away a variety of trivial uses of common FS which many have higher LPR due to the lexical and syntactic specificity of the FS. For instance, in \figref{fig:example} if the node \ex{keep \gap under wraps}  is turned on and has a minLPR of 8, then, if the minLPR of a node such as \ex{keep \gap under wraps for} is 4, clear(t) will be 0.5. We can now define d_{0}(t) as:

\begin{displaymath}
d_0(t) = cover(t) \cdot clear(t)
\end{displaymath}

Note that the effect of both covering and clearing is to  increase explainedness for covered/cleared nodes, which indirectly encourages the model to turn an $n$-gram like \ex{keep \gap under wraps} on in order to explain nodes in its vicinity, even supposing its minLPR were relatively low.

@@

The third mechanism of node interaction involves $n$-grams which overlap in the corpus. Though there are some exceptions, it is generally the case that independent formulaic sequences do not overlap. For example, given that \ex{be keep \gap under} and \ex{keep \gap under wraps} often appear together, overlapping on the tokens \ex{keep \gap under}, we do not want both being selected as an FS, even in the case where both have high minLPR. To address this problem, we use a mechanism somewhat different than above: rather than increasing the explainedness of having turned-off nodes, we decrease the explainedness of the turned-on nodes, which serves to express confusion at having overlapping FS. The idea behind covering and clearing is to encourage the model to turn on nodes which effectively explain the presence of nodes above and/or below it in the lattice, but for overlaps a penalty is a more appropriate mechanism. Let $oc(x,y)$ refer to number of times the $n$-grams corresponding to nodes $x$ and $y$ overlapped in the corpus, and $o_1\dotts o_m$ refer to turned-on nodes which overlap with our target node $t$. We define $d_1$, the exponential term for the $E_1(t)$ as:

@@

The effect of overlaps on the explainedness is hyperbolic: small amounts of overlapping has little effect, but with significant overlap it becomes prohibitive, effectively forcing one of the nodes to turn off.

@@

The dependence of the explainedness of nodes on their neighbors as discussed in the preceding section effectively prohibit a global optimization of the lattice. Fortunately, though a majority of nodes in the lattice are connected to each other though some path, most of the effects of nodes on each other are relatively local, and effective local optimizations can be made tractable by applying some simple restrictions. The main optimization loop consists of iterations over the lattice until complete convergence (no changes in the final iteration). For each iteration over the main loop, each potentially active node is examined in order to evaluate whether its current status is optimal given the current state of the lattice. The order that we do this has an effect on the result: among the obvious options, good results are obtained when ordering nodes by frequency, which gives an implicit advantage to relatively common $n$-grams, since they will tend to be turned on first.

@@

Given the relationships between nodes that we have defined above, it is obviously not sufficient to consider switching only the present node; If, for instance, one or more of \ex{be keep \gap under wraps}, \ex{under wraps}, or \ex{be keep \gap under} has been turned on, the covering, blocking, or overlapping effects of these other nodes will likely prevent a competing node like \ex{keep \gap under wraps} from being correctly activated. Instead, the algorithm identifies a small set of nodes (relevant nodes) which are the most important to status of the node under consideration whose status should be considered simultaneously. Since turned-off nodes have no direct effect on each other, only turned-on nodes above, below, or overlapping with the current node need be considered.  Once the relevant nodes have been identified, all nodes (including turned-off nodes) whose explainedness is affected by one or more of the relevant nodes are identified, and then a greedy search is carried out for the optimal configuration of the relevant nodes starting from an `all-on' state.

@@

\begin{itemize}
\item The complexity of the algorithm increases exponentially with the number of relevant nodes, so we limit the total number to 5. When there are more than 5 nodes turned on in the vicinity of the target node, relevant nodes are selected by ranking in terms of the influence of the nodes on each other as given by the relative change in their explainedness functions.
\item To avoid having to deal with storing and processing trivial overlaps, we exclude overlaps with a count of less than 5 from our lattice.
\item Many nodes have a minLPR which is a little larger than 1. There is very little chance these nodes will be activated by the algorithm, and so after applying hard covering we do not consider activating nodes with a minLPR of less than 2.
\end{itemize}

@@

	Lattice \emph{-ovr} & 0.52& \bf{0.51} & 0.51 & & 0.34& \bf{0.60} & 0.44& & 0.36 & \bf{0.67} & 0.47 & & 0.69 & \bf{0.51} & \bf{0.58} \\

				Lattice & \bf{0.67} & 0.43 & \bf{0.52} & & \bf{0.40}& 0.60 & \bf{0.48} & &\bf{0.44} & 0.56 & \bf{0.49} & & \bf{0.87} & 0.39 & 0.53 \\ 
            \hline
           \hline

@@

       \hline
			& P & R & F\\
			 \hline
			PMIrank & 0.20 & 0.29 & 0.23 \\
			minLPRrank & 0.34 & 0.45 & 0.39 \\
			LPR-seg & 0.42 & 0.45 & 0.43 \\
			LocalMaxs & \bf{0.56} & 0.39 & 0.46 \\
			DP-seg & 0.35 & \bf{0.71} & 0.47 \\
			Lattice & 0.47 & 0.63 & \bf{0.54} \\
       \hline
 \end{tabular}

@@

Our primary comparison is with the heuristic LPR model from Brooke et \al \shortcite{Brooke15b}, which is scalable to large corpora and includes gapped $n$-grams. For the BNC, we also compare, separately, the DP-seg model from Newman et \al \shortcite{Newman12} with recommended settings, and the LocalMaxs algorithm of Silva and Lopes \shortcite{Silva99} using the SCP as the association measure; because these other approaches only generate sequential multiword units, we use only the sequential part of the BNC test set. All comparison approaches have themselves been compared against a wide range of association measures in earlier work;\footnote{One popular association measure which is theoretically extensible to $n$-grams of any length but which has not been addressed in the context of unrestricted MWE/FS identification is the (log-)likelihood ratio \cite{Dunning93}. Calculating a likelihood ratio in the general case involves considering all possible ways an $n$-gram could have been generated statistically from its component words: that is, whereas minLPR only considers the addition of one word to each $n-1$-gram, likelihood ratios must consider all combinations of all component $n-1,2,3\ldots$-grams, including, in our case, not only contiguous but also non-contiguous components. The corresponding exponential increase in complexity renders it and other such methods impractical for unrestricted FS identification.} we do not repeat all these comparisons here, but we do consider a lexicon built from ranking $n$-grams according to the measure used in our lattice (minLPR) as well as PMI. For each of these two association measures we build a lexicon equal to the size of the lexicon produced by our model. All comparisons here are subject to the same $n$-gram frequency threshold as the main model, since the content of our test sets are also linked to this threshold.

@@

The main results for FS acquisition across all four corpora are shown in \tabref{tab:main}. As noted in previous work, simple statistical association measures like PMI do fairly poorly when faced with unrestricted $n$-grams of variable length: minLPR is clearly a much better statistic for this purpose. The LPRseg method of Brooke et \al \shortcite{Brooke15b} consistently outperforms simple ranking, and the lattice method proposed here does better still, with a margin that is fairly consistent across languages. Generally, using clearing and overlap node interactions offer an relatively large increase in precision at the cost of a smaller drop in recall, though the change is fairly symmetrical in Croatian. The Japanese and ICWSM corpus had relatively high precision and low recall, whereas both the BNC and Croatian corpus have low precision and high recall.

In the continuous FS test set for the BNC (\tabref{tab:BNC}), we found that both the LocalMaxs algorithm and the DP-seg method of Newman \cite{Newman12} were able to beat our other baseline methods with roughly similar f-scores, though both are well below our Lattice method. Some of the difference seems attributable to fairly severe precision/recall imbalance, though we unable to improve the f-score by changing the parameters from recommended settings for either model.

@@

Though the results across the four corpora are reasonably similar with respect to overall f-score, there are a significant discrepancies. By using the UniDict morpheme representation as the starting unit for Japanese, the model is, in essence, doing an extra layer of FS identification, one which is provided by word boundaries in the other languages. The result is that there are many more FS for Japanese. Precision is very high, and recall is relatively low. Under these conditions, methods which tend to increase precision at the cost of recall will be less desirable with respect to f-score, which explains why the full model is not the best in that one case.  Importantly, the initial $n$-gram statistics derived from corpus reflect that Japanese is different: the number of $n$-gram types, over length 4, for instance, is almost twice the number in the ICWSM, or almost 14\% of all $n$-gram types. One idea for future work is to automatically adapt to the particular properties of the input language/corpus in order to ensure a good balance between precision and recall.

At the opposite end of the precision/recall curve, the low precision of the BNC is almost certainly at least somewhat a reflection of its relatively small size: whereas the $n$-gram threshold we have used here results in a minimum counts of in the order of about 100 for the other three corpora used here, the BNC statistics includes $n$-grams which appear less than 10 times. This might be resolved, of course, by increasing the $n$-gram threshold, or simply avoiding small corpora altogether, but it may be useful to focus effort on how to building comprehensive FS lexicons even in relatively low resource situations. One idea we are actively pursing is modifying the calculation of the LPR metric to integrating uncertainty due to low counts.

There is one another more general explanation for the precision/recall imbalances we see across the four corpora: Both the BNC and the Croatian corpus are composed primarily (though not exclusively) of published texts written by professional writers, particularly news. The ICWSM and the Japanese corpus, on the other hand, are mostly blogs. Though all genres have FS, it is our observation that there is much more diversity in less controlled genres: many English gapped expressions, in particular, are used almost exclusively in relatively informal genres. The fact that last than 20\% of the FS derived from the BNC appears in the lexicon for the ICWSM is fairly telling of how important genre differences are to the range of FS that will be seen in a particular corpus.
